### 2020-10-12 input/background.tex
meow
better integration of def path

\begin{definition}[Logical partition]
  A logical \underline{p}artition $P_s$ originated at a
  \underline{s}ource process $s \in V$ is a set of interconnected
  processes that do not belong to another logical partition. The set
  of sources is $S$.
\end{definition}

\begin{definition}[Optimal partitioning]
  The optimal partitioning $\mathcal{P}$ is a set of logical
  partitions such that each process belongs to a logical partition
  comprising its closest source process. $\forall p \in P_{s_1\in S},
  \nexists P_{s_2 \in S}$ such that \TODO{$distance(p, s_2) <
    distance(p, s_1)$}.
\end{definition}

When $distance$ involves latency or number of hops between sources and
processes, the optimal partitioning consists in processing the
multiple destinations shortest path~\REF at each process.

\begin{definition}[Shortest path]
  A \underline{p}ath $\pi_{ij}$ from Process $i$ to Process $j$ is a
  sequence of vertices $[i, k_1, k_2, \ldots k_n, j]$ with $\forall
  0\leq m \leq n, \langle \pi_{ij}[m], \pi_{ij}[m+1] \rangle \in
  E$. Shortest \underline{p}aths $\Pi_{ij}$ are sets of paths such
  that there exists no lower sum of weights from $i$ to $j$.
\end{definition}

\TODO{Define operation alpha. Receipt and deliveries.}

Without global knowledge, processes only rely on their own behavior to
deduce the behavior of their neighbor processes.

\begin{definition}[\label{def:pathdelivery}\TODO{Shortest path} delivery]
  \TODO{Shortest path} delivery ensures the system eventually converges to
  optimal partitioning:
  \begin{inparaenum}[(i)]
  \item (\textbf{termination}) processes do not need to propagate
    operations they do not deliver: $\forall \langle i, j \rangle \in
    E, \neg d_i(\alpha_{s}^w)\implies \neg
    d_j(\alpha_{s}^{w+w_{ij}})$;
  \item (\textbf{propagation}) operations delivered by a process may
    also benefit neighbor processes: $\forall \langle i, j \rangle \in
    E, d_i(\alpha_{s}^{w_1}) \implies
    d_j(\alpha_{s}^{w_1+w_{ij}}) \vee
    d_j(\alpha_{*}^{w_2 < w_1 + w_{ij}})$.
  \end{inparaenum}
\end{definition}

@@

Algorithm~\ref{algo:add} shows the instructions that implement the
rules of Definition~\ref{def:pathdelivery} to converge towards optimal
partitioning. Figure~\ref{fig:add} illustrates its behavior on a
system comprising 4 processes $a$, $b$, $c$, and $d$. Two processes
$a$ and $c$ concurrently create a partition. They initialize their own
state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, Process $b$
receives $\alpha_{c}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. Process $b$
discards $\alpha_{b}^{1.5}$, for it does not improve its partition
distance. Processes $c$ and $d$ will never hear of Process $a$'s
partition. At scale, it saves bandwidth by limiting generated traffic
to subsets of processes. In Figure~\ref{fig:addD}, processes discard
last transiting messages. The system converged to the optimal
partitioning. Interestingly, while the receipt order of messages does
not impact on local control information, it impacts on control
information broadcast in the network. \TODO{explain with r() d() b().}

Transitive relationships ensure that each process gets its closest
source while accumulation of weights ensures that messages propagation
terminates.  Removals of partitions are more
challenging. \TODO{Figure.} \TODO{Also easy to solve when all
  processes receive all messages}. One could solve this issue by
removing partitions that were not advertised for a defined
time~\REF. However, relying on physical timeout could lead to either
premature removals of partitions when they are still operating; or
slow convergence where processes believe they belong to a partition
that was removed.

@@

Order on the broadcast.

>>>>

\begin{definition}[Logical partition]
  A logical \underline{p}artition $P_s$ originated at a
  \underline{s}ource process $s \in V$ is a set of interconnected
  processes that do not belong to another logical partition. The set
  of sources is $S \subseteq V$.
\end{definition}

Processes belong to logical partitions depending on path lengths. When
weights are latency or hop numbers, the optimal partitioning consists
in processing the multiple destinations shortest path at each process
\REF.

\begin{definition}[Shortest path]
  A \underline{p}ath $\pi_{ij}$ from Process $i$ to Process $j$ is a
  sequence of vertices $[i, k_1, k_2, \ldots k_n, j]$ with $\forall
  0\leq m \leq n, \langle \pi_{ij}[m], \pi_{ij}[m+1] \rangle \in
  E$. Shortest \underline{p}aths $\Pi_{ij}$ are sets of paths such
  that there exists no lower sum of weights $|\Pi_{ij}|$ from $i$ to
  $j$.
\end{definition}

\begin{definition}[Optimal partitioning]
  The optimal partitioning $\mathcal{P}$ is a set of logical
  partitions such that each process belongs to a logical partition
  comprising its closest source process.  $\forall p \in P_{s_1\in S},
  \nexists P_{s_2 \in S}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

Processes use reliable asynchronous message passing to reach optimal
partitioning together. Process $s$ generates a message
$\alpha_s^0$. Upon \underline{r}eceipt $r_i(\alpha_s^{w})$, Process
$i$ knows that Process $s$ created a partition at a path length of
$w$. Process $i$ may choose to \underline{d}eliver $d_i(\alpha_s^{w})$
if it improves its partitioning. Process $i$ may choose to propagate
its change through \underline{b}roadcast $b_i(\alpha_s^w)$.  Without
global knowledge, processes only rely on their own behavior to deduce
the behavior of their neighbor processes.

\begin{definition}[\label{def:transitivedelivery}Transitive delivery]
  Transitive delivery ensures the system eventually converges to
  optimal partitioning:
  \begin{inparaenum}[(i)]
  \item (\textbf{termination}) processes do not need to propagate
    operations they do not deliver: $\forall \langle i, j \rangle \in
    E, \neg d_i(\alpha_{s}^w)\implies \neg
    d_j(\alpha_{s}^{w+w_{ij}})$;
  \item (\textbf{propagation}) operations delivered by a process may
    also benefit neighbor processes: $\forall \langle i, j \rangle \in
    E, d_i(\alpha_{s}^{w_1}) \implies
    d_j(\alpha_{s}^{w_1+w_{ij}}) \vee
    d_j(\alpha_{*}^{w_2 < w_1 + w_{ij}})$.
  \end{inparaenum}
\end{definition}

@@

Algorithm~\ref{algo:add} shows the instructions that implement the
rules of Definition~\ref{def:transitivedelivery} to converge towards
optimal partitioning. Figure~\ref{fig:add} illustrates its behavior on
a system comprising 4 processes $a$, $b$, $c$, and $d$. Two processes
$a$ and $c$ concurrently create a partition. They initialize their own
state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, Process $b$
receives $\alpha_{c}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. Process $b$
discards $\alpha_{b}^{1.5}$, for it does not improve its partition
distance. Processes $c$ and $d$ will never hear of Process $a$'s
partition. At scale, it saves bandwidth by limiting generated traffic
to subsets of processes. In Figure~\ref{fig:addD}, processes discard
last transiting messages. The system converged to the optimal
partitioning.  Transitive relationships ensure that each process gets
its closest source while accumulation of weights ensures that messages
propagation terminates.  Interestingly, while the receipt order of
messages does not impact on local control information
($r_b(\alpha_a^{1.5}) \cdot r_b(\alpha_a^{1}) \Leftrightarrow
r_b(\alpha_a^{1}) \cdot r_b(\alpha_a^{1.5}) \implies
\TODO{d_b(\alpha_a^{1})}$), it impacts on control information
broadcast in the network ($r_b(\alpha_a^{1.5}) \cdot r_b(\alpha_a^{1})
\implies b_b(\alpha_{b}^{1.5}) \cdot b_b(\alpha_{b}^{1})$ while
$r_b(\alpha_a^{1}) \cdot r_b(\alpha_a^{1.5}) \implies
b_b(\alpha_a^{1})$).

While only adding logical partitions to the distributed system is
straightforward, removing them introduces additional complexity.

\begin{definition}[Dynamic optimal partitioning]
  \TODO{add and del operations.}
\end{definition}

\TODO{Figure.} \TODO{Also easy to solve when all
  processes receive all messages}. One could solve this issue by
removing partitions that were not advertised for a defined
time~\REF. However, relying on physical timeout could lead to either
premature removals of partitions when they are still operating; or
slow convergence where processes believe they belong to a partition
that was removed.

@@
